---
title: "Webinar-3"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cell to load the tidyverse library:
```{r}
library(tidyverse)
```

#### Execute the following cell to load the data. Refer to this website http://openmv.net/info/food-texture for data description:
```{r}
file = 'http://openmv.net/file/food-texture.csv'
#file = 'Data/food-texture.csv'
foodData = read.csv(file, header = TRUE, row.names = 1)
```

#### Print structure of dataframe
```{r}
str(foodData)
```

#### Print first 5 samples of data frame
```{r}
head(foodData, n = 5)
```

#### Get the number of samples and features
```{r}
nrow(foodData)
ncol(foodData)
```

#### Rename Oil column to OilPercentage
```{r}
foodData = foodData %>% rename(OilPercentage = Oil)
str(foodData)
```

#### In-built functions for dataframes
```{r}
# Mean oil percentage across all samples
mean(foodData$OilPercentage)

# Mean-centering of OilPercentage
foodData$OilPercentageMC = foodData$OilPercentage - mean(foodData$OilPercentage)
head(foodData)

# Sum of the squared deviation from the mean
sum((foodData$OilPercentageMC)^2)

# Average of the squared deviation from the mean
(1/(nrow(foodData)-1))*sum((foodData$OilPercentageMC)^2) # also called the variance

# Variance of OilPercentage
var(foodData$OilPercentage)

# Standard deviation of OilPercentage
sd(foodData$OilPercentage)
```

#### Visualize the OilPercentage features
```{r}
bins = seq(mean(foodData$OilPercentage)-3*sd(foodData$OilPercentage), mean(foodData$OilPercentage)+3*sd(foodData$OilPercentage), by = 1.0)
p1 = ggplot(data = foodData) + 
  geom_histogram(aes(x = OilPercentage, y = ..count..),
                 breaks = bins, colour = 'black',
                 fill = 'steelblue') +
  labs(x = 'Oil Percentage', y = 'Frequency') +
  ggtitle('Histogram of Oil Percentage Values')
p1
```

#### Modify crispy column to reflect high (0) or low (1) crispiness
```{r}
foodData = foodData %>% mutate(Crispy = ifelse(Crispy > 11, 'high', 'low'))
head(foodData)
str(foodData)
```

##### Change Crispy column to factor (categorical) type
```{r}
#factor(foodData$Crispy)
foodData['Crispy'] = lapply(foodData['Crispy'], factor)
str(foodData)
```

#### Scatter plot between OilPercentage and Density
```{r}
p2 = ggplot(data = foodData) +
  geom_point(aes(x = Density, y = OilPercentage, color = Crispy), size = 1)
p2
```

#### Scale continuous columns of the dataframe
```{r}
ind_continuous = sapply(foodData, is.numeric)
foodData[ind_continuous] = lapply(foodData[ind_continuous], scale)
head(foodData)
```

#### Calculate the correlation matrix for the continuous features
```{r}
cor(foodData[ind_continuous])
```

#### Scatter plot between Density and Hardness
```{r}
p3 = ggplot(data = foodData) +
  geom_point(aes(x = Density, y = Hardness), size = 1)
p3
```

#### Scatter plot between Density and Fracture
```{r}
p2 = ggplot(data = foodData) +
  geom_point(aes(x = Density, y = Fracture), size = 1)
p2
```